Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2022-02-04 src/ms.tex
abstract revision

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{\today}
\revised{\today}
%\accepted{\today}

@@

\begin{abstract}
We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic parameters from the California-Kepler Survey and LAMOST survey to demonstrate the existence of a pile-up at the long-period edge of the effective temperature--rotation period distribution of main-sequence stars with temperatures exceeding $\sim$\teffmin. The long-period pile-up was previously predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which wind-driven angular momentum losses cease once stars reach a critical Rossby number. Stars in the long-period pile-up are found to have a wide range of ages (\agerange), meaning that, along the pile-up, rotation period is strongly predictive of a star's surface temperature but only weakly predictive of its age. In the effective temperature range of 5000--6250~K the long-period edge is well-described by a curve of constant Rossby number, with a critical value of \rocritfinal. These results imply that surface rotation periods have limited predictive power for the age determination of F- and early G-type stars older than \agemin. Using kernel density estimation, we additionally detect a secondary pile-up associated with the short-period edge of the joint temperature–period distribution, which is also well-described by a curve of constant Rossby number. The short-period pile-up is not a prediction of the weakened magnetic braking hypothesis, but may instead be related to a phase of slowed surface spin-down due to core–envelope coupling. The same mechanism was previously proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters. The relative dearth of stars with intermediate rotation periods between the short- and long-period pile-ups is also well-described by a curve of constant Rossby number, which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars. These observations provide further support for the hypothesis that the period gap is due to stellar astrophysics, rather than a non-uniform star-formation history in the Kepler field.
\end{abstract}

>>>>

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{February 4, 2022}
\revised{\today}
%\accepted{\today}

@@

\begin{abstract}
We combine stellar surface rotation periods determined from NASA's Kepler mission with spectroscopic temperatures to demonstrate the existence of pile-ups at the long-period and short-period edges of the temperature--period distribution for main-sequence stars with temperatures exceeding $\sim$\teffmin. The long-period pile-up is well-described by a curve of constant Rossby number, with a critical value of \rocritfinal. The long-period pile-up was predicted by \citet{vanSaders2019} as a consequence of weakened magnetic braking, in which wind-driven angular momentum losses cease once stars reach a critical Rossby number. Stars in the long-period pile-up are found to have a wide range of ages (\agerange), meaning that, along the pile-up, rotation period is strongly predictive of a star's surface temperature but weakly predictive of its age. The short-period pile-up, which is also well-described by a curve of constant Rossby number, is not a prediction of the weakened magnetic braking hypothesis but may instead be related to a phase of slowed surface spin-down due to core–envelope coupling. The same mechanism was proposed by \citet{Curtis2020} to explain the overlapping rotation sequences of low-mass members of differently aged open clusters. The relative dearth of stars with intermediate rotation periods between the short- and long-period pile-ups is also well-described by a curve of constant Rossby number, which aligns with the period gap initially discovered by \citet{McQuillan2013b} in M-type stars. These observations provide further support for the hypothesis that the period gap is due to stellar astrophysics, rather than a non-uniform star-formation history in the Kepler field.
\end{abstract}
